CF Registry Analysis

### 2024-02-16 Christine Chan/CF Registry/cf_registry_analysis.qmd
forest plot with mostly base R

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "bmi_perc_cat", "MutClass", "Race",
  "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
  "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_mycocultureresults3", "A_pulmonarycomplications1", "A_corticosteroids1",
  "cfrd_status_annual", "A_Vx770", "A_VX809comb", "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Race = "Race", Hispanicrace = "Hispanic",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  ppFEV1_cat = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc_cat = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_pulmonarycomplications1 = "ABPA",
  A_mycocultureresults3 = "Microorganisms in myco culture",
  eDWID = "ID",birth_year="Birth year",tstart="Interval start",tstop="Interval stop",age_at_entry="Age at entry",Age_YrEnd="Age",bmi_perc = "BMI %ile"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd,
    Age_YrEnd >= 10
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(
    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
    age_at_entry, everything()
  )
# Convert from tibble to dataframe for survival packages
s <- as.data.frame(s)
# CFRD status to numeric
s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
# Drop under 10s
s$age_group <- droplevels(s$age_group)
# Labels
label(s) = labels[colnames(s)]
```

@@

```{r}
# Formula for base model
x <- c(
  "age_group", "Gender", "Race","Hispanicrace", "bmi_perc_cat", "MutClass", 
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat", 
  "NumPulmExacerbation", "A_pulmonarycomplications1", "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, cfrd_status_annual)"
f <- as.formula(paste0(y, "~", paste0(x, collapse = "+")))
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
ggforest(cox_res, data = s)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph = test.ph[rownames(test.ph) != "GLOBAL",]
# Find which variables technically violate PH assumption
time_varying = rownames(test.ph)[test.ph$p < 0.05]
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```

>>>>

```{r}
forest_plot <- function(mod, time_varying = NULL, labels = labels) {
  coeff <- data.frame(summary(mod)$coefficients)
  coeff$ll <- coeff$coef - (1.96 * coeff$se.coef.)
  coeff$ul <- coeff$coef + (1.96 * coeff$se.coef.)
  # Clean up variable names
  l <- unlist(sapply(labels, paste0, ": "))
  coeff$names <- rownames(coeff)
  coeff$names <- str_replace_all(coeff$names, l)
  coeff$names <- ordered(coeff$names, levels = unique(coeff$names))
  # Exponentiate
  coeff$exp.ll <- exp(coeff$ll)
  coeff$exp.ul <- exp(coeff$ul)
  # Plot
  p <- ggplot(coeff, aes(y = fct_rev(names))) +
    geom_point(aes(x = exp.coef.), shape = 15, size = 3) +
    geom_linerange(aes(xmin = exp.ll, xmax = exp.ul)) +
    geom_vline(xintercept = 1, linetype = "dashed") +
    xlab("Hazard Ratio") +
    theme_classic() +
    theme(axis.ticks.y = element_blank(), axis.title.y = element_blank())
}
```

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "bmi_perc_cat", "MutClass", "Race",
  "Hispanicrace", "A_IsOnEnzymes", "ppFEV1_cat", "tube_feeding",
  "NumPulmExacerbation", "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_mycocultureresults3", "A_pulmonarycomplications1", "A_corticosteroids1",
  "cfrd_status_annual", "A_Vx770", "A_VX809comb", "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Race = "Race", Hispanicrace = "Hispanic",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  ppFEV1_cat = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc_cat = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_pulmonarycomplications1 = "ABPA",
  A_mycocultureresults3 = "Microorganisms in myco culture",
  eDWID = "ID", birth_year = "Birth year", tstart = "Interval start", tstop = "Interval stop", age_at_entry = "Age at entry", Age_YrEnd = "Age", bmi_perc = "BMI %ile"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd,
    Age_YrEnd >= 10
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(
    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
    age_at_entry, everything()
  )
# Convert from tibble to dataframe for survival packages
s <- as.data.frame(s)
# CFRD status to numeric
s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
# Drop under 10s
s$age_group <- droplevels(s$age_group)
# Labels
label(s) <- labels[colnames(s)]
```

@@

```{r}
# Formula for base model
x <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_pulmonarycomplications1", "A_corticosteroids1"
)
y <- "Surv(tstart, tstop, cfrd_status_annual)"
f <- as.formula(paste0(y, "~", paste0(x, collapse = "+")))
cox_res <- coxph(f, data = s, cluster = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
ggforest(cox_res, data = s)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
# ggcoxzph(test.ph)
# ggcoxdiagnostics(cox_res,type = "martingale")
# ggcoxdiagnostics(cox_res,type = "deviance")
# ggcoxdiagnostics(cox_res,type = "score")
# ggcoxdiagnostics(cox_res,type = "schoenfeld")
```
